FATE summary results

### 2019-12-11 paper draft.Rmd
Updated plots and loadings threshold

```{r, echo=F}
GOA.biol <- readRDS("GOA_biol_9.9.19.rds") 
GOA.biol$summary
```

@@

@@

plot.a <- ggplot(plot.load, aes(key, value)) +
  theme_bw() +
  geom_violin(fill=cb[3]) +
  geom_hline(yintercept = 0) +
  theme(axis.title.y = element_blank()) + ylab("Loading") +
  coord_flip() +
  ggtitle("a) Loadings")
```

@@

```{r, include=F}
# first, define the function
trend_lm <- function(rotated_modelfit,
  all_years, selected_years,
  climate_dat = NULL,
  trend_number = 1, samples = 50,
  ITER = 1000, CHAINS = 1, n_eff = 50, rhat = 1.1, show_plot = TRUE) {

@@

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2017] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

@@

names(mod)[2:4] <- c("1972-1988", "1989-2013", "2014-2017")
summary(mod)

@@

melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2017"))
names(melt.mod)[3] <- "era"
melt.mod$covar <- "SST"

@@

pdo <- read.csv("pdo.csv")

win.pdo <- pdo$NDJFM

years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2017

# set up climate data
cli_dat <- data.frame(Year=pdo$YEAR, PDO=win.pdo)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(PDO = arm::rescale(PDO))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2017] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + PDO:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2017,
  selected_years = 1972:2016, climate_dat=cli_dat)

names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2017")

summary(mod2)

for(i in 2:4){
x <- sort(mod2[,i])
print(names(mod2)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

melt.mod2 <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2017"))
names(melt.mod2)[3] <- "era"
melt.mod2$covar <- "PDO"

@@

# make a SST-only plot
ggplot(melt.mod, aes(x=value, fill=era)) +
  theme_bw() + 
  geom_density(alpha=0.6) +
  geom_vline(xintercept = 0, lty=2) +
  scale_fill_manual(values = cb[c(4, 6, 2)]) +
  theme(legend.title = element_blank(), legend.position = c(0.8, 0.8)) +
  xlab("Slope") +
  xlim(-5,12)

>>>>

```{r, echo=F}
GOA.biol <- readRDS("GOA_biol_11.4.19.rds") 
GOA.biol$summary
```

@@

prop1 <- apply(loadings.1, 2, f)
prop1
keep <- prop1 > 0.9 # changed threshold to 0.9

@@

plot.a <- ggplot(plot.load, aes(key, value)) +
  theme_bw() +
  geom_violin(fill=cb[3], color=NA) +
  geom_hline(yintercept = 0) +
  theme(axis.title.y = element_blank()) + ylab("Loading") +
  coord_flip() + ylim(-1.5, 1.5) +
  ggtitle("a) Loadings")
```

```{r, echo=F}
# trends
plot.trend <- data.frame(year=1972:2019, trend=as.vector(rotated$trends_mean),
                         lo=as.vector(rotated$trends_lower),
                         hi=as.vector(rotated$trends_upper))

png("combined GOA biology loading and trend plot.png", 8, 4, units = "in", res=300)

@@

```{r, include=F}
# first, define the function
trend_lm <- function(rotated_modelfit,
  all_years, selected_years,
  climate_dat = NULL,
  trend_number = 1, samples = 500,
  ITER = 1000, CHAINS = 1, n_eff = 50, rhat = 1.1, show_plot = TRUE) {

@@

@@

names(mod)[2:4] <- c("1972-1988", "1989-2013", "2014-2019")
summary(mod)

@@

melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2019"))
names(melt.mod)[3] <- "era"
melt.mod$covar <- "SST"

@@

```{r, echo=F}
# 
# pdo <- read.csv("pdo.csv")
# 
# win.pdo <- pdo$NDJFM
# 
# years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2017
# 
# # set up climate data
# cli_dat <- data.frame(Year=pdo$YEAR, PDO=win.pdo)
# 
# cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(PDO = arm::rescale(PDO))
# 
# # cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
# 
# cli_dat$slope <- 1
# cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
# cli_dat$slope[cli_dat$Year %in% 2014:2017] <- 3
# cli_dat$slope <- as.factor(cli_dat$slope)
# 
# # rename rotated
# r <- rotated
# 
# Y = rnorm(length(cli_dat$slope))
# cli_dat = as.data.frame(model.matrix(Y ~ Year + PDO:slope, data=cli_dat))
# # replace ':' in name for formula
# names(cli_dat) = gsub(":","",names(cli_dat))
# # remove named intercept
# cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
# 
# set.seed(99)
# mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2017,
#   selected_years = 1972:2016, climate_dat=cli_dat)
# 
# names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2017")
# 
# summary(mod2)
# 
# # get 95% CIs
# 
# for(i in 2:4){
# x <- sort(mod2[,i])
# print(names(mod2)[i])
# print(paste("median: ", median(x)))
# print(paste("LCI: ", x[(0.025*25000)]))
# print(paste("UCI: ", x[(0.975*25000)]))
# }
# 
# 
# melt.mod2 <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2017"))
# names(melt.mod2)[3] <- "era"
# melt.mod2$covar <- "PDO"

@@

# make a SST-only plot
ggplot(melt.mod, aes(x=value, fill=era)) +
  theme_bw() + 
  geom_density(alpha=0.6) +
  geom_vline(xintercept = 0, lty=2) +
  scale_fill_manual(values = cb[c(4, 6, 2)]) +
  theme(legend.title = element_blank(), legend.position = c(0.8, 0.8)) +
  xlab("Slope") + ylab("Density") +
  xlim(-5,12)
